Thermodynamic Structure of the Solar Corona: Tomographic Reconstructions and MHD Modeling

### 2020-03-08 response_to_review_SOLA-D-20-00015_v1.txt

Review of Thermodynamic Structure of the Solar corona: tomographic reconstructions and MHD modeling

This is a well-structured and clearly-written paper, presenting a detailed comparison of tomographic reconstructions of the low solar corona (which gives density and temperature), and a global MHD model. The work is important and valuable to the field, and is suitable for publication in Solar Physics. There are no major scientific concerns. The paper should be ready for publication if the following, mostly minor, concerns are addressed:

Medium concerns
-----------------------------------------------------------------------------------------
1. All of the multi-panel figures that show longitude-latitude maps (e.g. figure 1, 2, 3 etc) are too small, it is very difficult to see any details. There is a lot of white space between the panels that can be removed, and this would be one way to instantly make the maps larger.

We agree. We improved the use of space in Figures 1,2,8 and 9.

@@

We agree. Actually this affects all following Figures #: 3, 4, 11, and 14. In those, we applied the following changes: Red->Orange and Magenta->Red. These emulate our previous choice of colors in Nuevo et al. (2013) and Lloveras et al. (2017). We like to keep all "up" structures in a "warm" palette. We edited text accordingly all along the manuscript. We believe the colors are now clearly distinguishable in all figures.

-----------------------------------------------------------------------------------------
3. Page 14, paragraph starting at line 34. You discuss the latitudinal distribution of short and long loops, and state that the small legs mainly populate the equatorial latitudes, and longer legs at higher latitudes. But this is a skewed result - their spatial positioning was a criteria for their categorisation, so this should not be presented as a scientific result.

Thanks for pointing this out.... ESTO REQUIERE CAMBIOS EN EL PAPER JUSTIFICANDO CON CLARIDAD LA ELECCION DE LOS LIMITES LATITUDINALES Y LA SELECCION DE SOLO UP EN TYPE II Y III. Y UNA RESPUESTA ACORDE.

That's right, we used it as a criterion of selection aiming to distinguish different layers. we changed the text avoiding confusion

@@

DEFINAMOS EL CAMPO ESCALAR Xi = (X_MHD - X_tom) / X_tom, DONDE "X" ES CUALQUIER RESULTADO A LO LARGO DE LOOP (Ncb, LAMDA_n, ó <Tm>). LA TABLA 2 INFORMA LOS VALORES: "[Med(X_MHD) - Med(X_tom)]/Med(X_tom)".

PROPONEMOS DEJAR LOS HISTOGRAMAS DE LAS FIGS 12 Y 13 COMO ESTÁN, PERO EXPANDIR LA TABLA 2 A QUE INCLUYA DOS COLUMNAS MÁS INDICANDO ESTOS DOS INDICADORES DE CADA CAMPO "Xi":
* Med(Xi)  (debería aproximarse a los porcentuales y signos entre paréntesis que actualmente informa la Tabla 2).
* MAD = Med(Abs(Xi-Med(Xi))).

LUEGO EDITAMOS TABLA Y TEXTO DEL PAPER.

The populations shown in the histograms are about types of legs associated with different latitudinal regions. The selection of these legs is not the same for AWSoM and DEMT. This is why calculating the MAD a on the relationship of the populations of the histograms reported is not simple. In the future we plan to make a comparative analysis leg-to-leg between AWSoM and DEMT and in that case we will undoubtedly report some statistics associated with the standard deviation such as being the MAD.

@@

-----------------------------------------------------------------------------------------
5. You discuss the less reliable results from AIA in coronal holes. The very recent paper by Saqri et al (Solar Physics 295, 2020) is relevant here, and should be referenced.

ESTO FUE SOLUCIONADO. EL ORIGEN DEL PROBLEMA NO RADICA EN STRAY-LIGHT DECON. EL EFECTO DE ESTO EN TOM, APLICADO A DATOS EUVI, FUE EXPLORADO EN LLOVERAS ET AL. (2017). CIERTAMENTE NOS INTERESA INCORPORAR ESTO AL PIPELINE DE AIA SI LA PFS ESTÁ BIEN CARACTERIZADA, PERO ES UN TEMA QUE DEBEMOS INVESTIGAR. EXPLICAR COMO SE SOLUCIONÓ Y REMOVER TODO LO NO NECESARIO EN EL PAPER (AZUL) Y AGREGAR COSAS NUEVAS (ROJO).

The problem in the magnetically open AIA region was actually because the minimization in the voxels used to converge to a local minimum. This was resolved by modifying the starting values of the minimization. Now the voxels takes correct values of density and temperature. This affects the figures that report type III legs for CR2208. Therefore, figures 4, 5, 10, 13 and 14 were modified. Now the temperature gradients are correct and the temperature profile reported in Figure 10 is improved.

### 2020-03-08 lloveras_2020_v2.tex
retoques del referi

{For both rotations, the top panels of Figure \ref{rpoint_demt} show the latitude-longitude location (at heliocentric height $r=1.105\,\mrsun$) of all traced field line legs for which criterion (i) of Section \ref{trace} is met. Open legs are indicated in gray color and closed ones in black color. \del{For each leg, the fits to tomografic temperature and density were applied, as given by Equations \ref{Nfit}-\ref{Tfit}.} Considering the DEMT data points and the resulting fits along each leg, the bottom panels of Figure \ref{rpoint_demt} show the latitude-longitude location of the subset for which also both criteria (ii) and (iii) of Section \ref{trace} are met. Using a four-color code, type 0, I, II and III legs are shown in blue, red, {magenta and cyan} color, respectively.} Of the {$\approx 44000$} legs selected for CR-2082, 20\% are type 0, 31\% are type I, 23\% are type II and 26\% type III. On the other hand, of the \edit{$\approx 50000$} legs selected for CR-2208, 10\% are type 0, \edit{38}\% are type I, \edit{27}\% are type II and \edit{25}\% type III.

@@

{Several details from Figure \ref{perf_lat} are \edit{worth highlighted}. Firstly, at most latitudes the overall agreement of the electron density of both models is within $\approx 20\%$ for CR-2082, and $\approx 5\%$ for CR-2208. The noticeable exception is to be found near the open/closed boundaries of both target rotations, where the disagreement between both models can be up to twice as much. In the case of the electron temperature, for both rotations the models agree within $\approx 15\%$ at all latitudes \del{of the streamer belt, showing larger discrepancies in the CHs}. Secondly, for both rotations, and for both models, these plots clearly show the relatively lower temperatures characterizing the low-latitudes of the equatorial streamer belt compared to its mid-latitudes. Thirdly, for both rotations, the latitude of the open/closed magnetic boundary in both hemispheres matches the location of the strongest latitudinal gradient of the DEMT electron density. Note this is not the case for the AWSoM model, that shows a minimum density at the open/closed boundary. \edit{Lastly, the DEMT} electron density decreases from the open/closed boundary towards the poles (in both hemispheres of the two rotations), while the AWSoM model shows the opposite trend. {For comparison,\edit{$B_r$ in the CHs increases} from the open/closed boundary towards the poles for CR-2082, while showing local maxima around latitudes $-75\mdeg$ and $+70\mdeg$ in the case of CR-2208.}}

@@

\del{Concerning the DEMT reconstructions, Figure \ref{rpoint_demt} shows that, for both rotations,} \edit{Magnetic} field lines of type 0, I and II are associated with increasingly outer {layers} of the equatorial streamer belt. \del{Figure \ref{histos_fulldemt} and Table \ref{tabla_demt} show in 3D quantitative detail how} \edit{These progressively} outer magnetic structures exhibit decreasing coronal base density, increasing density scale height, and increasing electron temperature. For both rotations we find that down legs populate the low latitudes of the streamer belt, while up legs dominate its mid-latitudes. \del{DIEGO: esto no es cierto, es nuestra definicion y no un resultado.} Also, in the case of CR-2082 the fraction of down legs is significantly larger than for CR-2208. These findings are consistent with previous studies by \citet{huang_2012} and \citet{nuevo_2013}. In the case of the latter, they include in their analysis target CR-2081, which is a rotation almost identical to our target CR-2082. Our results for target CR-2082 compare very well with those of \citet{nuevo_2015} and \citet{lloveras_2017} for target CR-2081. As our study uses the improved version of the DEMT technique, such comparison provided a consistency \edit{check.} \del{for the changes implemented in the DEMT technique.}

>>>>

For both rotations, the top panels of Figure \ref{rpoint_demt} show the latitude-longitude location (at heliocentric height $r=1.105\,\mrsun$) of all traced field line legs for which criterion (i) of Section \ref{trace} is met. Open legs are indicated in gray color and closed ones in black color. \del{For each leg, the fits to tomografic temperature and density were applied, as given by Equations \ref{Nfit}-\ref{Tfit}.} Considering the DEMT data points and the resulting fits along each leg, the \edit{mid} panels of Figure \ref{rpoint_demt} show the latitude-longitude location of the subset for which also both criteria (ii) and (iii) of Section \ref{trace} are met. Using a four-color code, type 0, I, II and III legs are shown in blue, \edit{orange, red and cyan} color, respectively. \edit{The bottom panels of same figure shows histograms of the latitudinal distribution of the legs using same color code.} Of the {$\approx 44000$} legs selected for CR-2082, 20\% are type 0, 31\% are type I, 23\% are type II and 26\% type III. On the other hand, of the \edit{$\approx 50000$} legs selected for CR-2208, 10\% are type 0, \edit{38}\% are type I, \edit{27}\% are type II and \edit{25}\% type III.

@@

{Several details from Figure \ref{perf_lat} are \edit{worth highlighted}. Firstly, at most latitudes the overall agreement of the electron density of both models is within $\approx 20\%$ for CR-2082, and $\approx 5\%$ for CR-2208. The noticeable exception is to be found near the open/closed boundaries of both target rotations, where the disagreement between both models can be up to twice as much. In the case of the electron temperature, for both rotations the models agree within $\approx 15\%$ at all latitudes \del{of the streamer belt, showing larger discrepancies in the CHs}. Secondly, for both rotations, and for both models, these plots clearly show the relatively lower temperatures characterizing the low-latitudes of the equatorial streamer belt compared to its mid-latitudes. Thirdly, for both rotations, the latitude of the open/closed magnetic boundary in both hemispheres matches the location of the strongest latitudinal gradient of the DEMT electron density. Note this is not the case for the AWSoM model, that shows a minimum density at the open/closed boundary. \edit{Lastly, the DEMT} electron density decreases from the open/closed boundary towards the poles (in both hemispheres of the two rotations), while the AWSoM model shows the opposite trend. {For comparison, \edit{$B_r$ in the CHs increases} from the open/closed boundary towards the poles for CR-2082, while showing local maxima around latitudes $-75\mdeg$ and $+70\mdeg$ in the case of CR-2208.}}

@@

\del{Concerning the DEMT reconstructions, Figure \ref{rpoint_demt} shows that, for both rotations,} \edit{Magnetic} field lines of type 0, I and II \edit{were selected to be associated} with increasingly outer {layers} of the equatorial streamer belt. \del{Figure \ref{histos_fulldemt} and Table \ref{tabla_demt} show in 3D quantitative detail how} \edit{These progressively} outer magnetic structures exhibit decreasing coronal base density, increasing density scale height, and increasing electron temperature. For both rotations we find that down legs populate the low latitudes of the streamer belt, while up legs dominate its mid-latitudes. Also, in the case of CR-2082 the fraction of down legs is significantly larger than for CR-2208. These findings are consistent with previous studies by \citet{huang_2012} and \citet{nuevo_2013}. In the case of the latter, they include in their analysis target CR-2081, which is a rotation almost identical to our target CR-2082. Our results for target CR-2082 compare very well with those of \citet{nuevo_2015} and \citet{lloveras_2017} for target CR-2081. As our study uses the improved version of the DEMT technique, such comparison provided a consistency \edit{check.} \del{for the changes implemented in the DEMT technique.}
